---
title: "HG002 PacBio CCS Small Variants"
author: "Nate Olson"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r load_packages, message = FALSE, echo = FALSE}
library(happyR)
library(tidyverse)
```


# Overview
Analysis of small variant callsets generated from PacBio CCS data. 
The long read - low error rate data can potentially correct mistakes in and expand the HG002 benchmark set.  

* datasets  
* variant callsets - Two SNV+indel callsets: GATK4 and DeepVariant against the GRCh37 reference.  
* benchmarking - Callsets benchmarked against the HG002 V3.3.2 benchmark callset using the precisionFDA app, vcfeval + hap.py with custom stratifications 

# Approach/ Methods
## Benchmarking 
- Stratified comparison to our current small variant benchmark (v3.3.2).  
- Identify where callsets perform well and poorly  
- Identify where callsets perform well with a high fraction of calls in a stratification outside the benchmark regions.  

## Manual Curation
A number of false positive and negative sites were randomly sampled to further evaluate the potential for using CCS callsets to correct errors in benchmark sets.
A total of 80 variants were randomly sampled for manual curation.  

- 10 from each set  
    - FP and FN  
    - SNP and Indels  
    - Target categories and other categories  

Target categories - lowcmp_SimpleRepeat_homopolymer_6to10, lowcmp_SimpleRepeat_homopolymer_gt10, and lowcmp_SimpleRepeat_imperfecthomopolymer_gt10_slop5

<!-- 
Especially for indels, most of the errors are in homopolymers, so it’s helpful to segregate these. 
For SNPs, it might also be useful to select 10 random from the most difficult to map regions (map_l250_m0_e0), unless most of the non-homopolymer SNPs are in these regions. 
Definitely let me know if your look at the data suggests a different strategy though! I can help with the manual curation when I return from AMP. 
-->

## Benchmark Region Expansion Estimate
Estimating of how many extra variants and regions CCS might help add to the GIAB benchmark. Perhaps, regions with good MQ and coverage excluding homopolymers, and the CCS variants in these regions.  

- Need to work out approach

# Results 

## Benchmarking  

<!-- - Subsets to look at  -->
<!--     - `*`   -->
<!-- 	- HG002allgenomespecific   -->
<!-- 	- HG002allgenomespecificandifficult   -->
<!-- 	- alldifficultregions   -->
<!-- 	- decoy   -->
<!-- 	- lowcmp_AllRepeats_51to200bp_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_AllRepeats_gt200bp_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_AllRepeats_gt95identity_slop5   -->
<!-- 	- lowcmp_AllRepeats_lt51bpTRs_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_AllRepeats_lt51bp_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_SimpleRepeat_homopolymer_6to10   -->
<!-- 	- lowcmp_SimpleRepeat_homopolymer_gt10   -->
<!-- 	- lowcmp_SimpleRepeat_imperfecthomopolymer_gt10_slop5   -->
<!-- 	- map_all   -->
<!-- 	- map_l250_m0_e0   -->
<!-- 	- notinHG002allgenomespecificandifficult   -->
<!-- 	- notinalldifficultregions   -->
<!-- 	- notinlowcmp_AllRepeats_gt95identity_slop5   -->
<!-- 	- segdup   -->

```{r message = FALSE}
## Loading data
hap_list <- paste0("data/happy_output/results/result_", c(1,2)) %>%
    map(read_happy) %>% 
    set_names(c("GATK4","DeepVariant"))

## Creating a tidy data frame
extend_df <- hap_list %>% 
    map("extended") %>% 
    bind_rows(.id = "query_method")

ext_trim_df <- extend_df %>% 
    ## Excluding non-HG002 genome specific stratifications
    filter(!str_detect(Subset, "HG00[1,3,4,5]")) %>% 
    ## excluding strat with < 1000 obs - lack of statistical power
    filter(Subset.Size > 1000) %>% 
    ## Subset must have at least one variant in region
    filter(Subset.IS_CONF.Size > 0)

## For initial set of plots
gatk_ext_df <-  filter(ext_trim_df, 
                       query_method == "GATK4", 
                       Subtype == "*")

gatk_metric_df <- gatk_ext_df %>% 
    group_by(Type, Subset) %>% 
    select(contains("METRIC")) %>% 
    gather("Metric", "Value", -Type, -Subset) 

```


Overall high precision and recall for SNPs, and indels not in difficult regions. 
DeepVariant performs consistently better than GATK4. 
The high frac_NA for SNPs in all difficult regions along with relatively high precision and recall indicates that using this in the GIAB integration pipeline may result in expanded benchmark regions.  
- Difficult regions with low indel recall and precision, specifically homopolymers and imperfect homopolymers 


```{r}
## Do we have all difficult regions not in HG002allgenomespecific
high_level_strat <- c("*",
                       "alldifficultregions",
                       "notinHG002allgenomespecificandifficult",
                       "notinalldifficultregions")

ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(Subset %in% high_level_strat) %>% 
    select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -45, hjust = 0))
```

```{r}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Subset %in% high_level_strat) %>% 
    filter(Filter == "PASS") %>% 
    rename(Callset = query_method,
            Recall = METRIC.Recall, 
           Precision = METRIC.Precision, 
           F1 = METRIC.F1_Score, 
           Frac_NA = METRIC.Frac_NA) %>%
    select(Callset, Type, Subset, Recall, Precision, F1, Frac_NA, Subset.Size) %>% 
    arrange(Type, Subset) %>% 
    knitr::kable(digits = 3, 
                 caption = "Metrics for high level stratifications.")
```

```{r, fig.cap = "GATK4 and DeepVariant precision v. recall for a subset of the stratifications."}
strat_of_interest <- c("*","HG002allgenomespecific",
                       "HG002allgenomespecificandifficult",
                       "alldifficultregions","decoy", 
                     "lowcmp_AllRepeats_51to200bp_gt95identity_merged_slop5",
                       "lowcmp_AllRepeats_gt200bp_gt95identity_merged_slop5",
                       "lowcmp_AllRepeats_gt95identity_slop5",
                    "lowcmp_AllRepeats_lt51bpTRs_gt95identity_merged_slop5",
                       "lowcmp_AllRepeats_lt51bp_gt95identity_merged_slop5",
                       "lowcmp_SimpleRepeat_homopolymer_6to10",
                       "lowcmp_SimpleRepeat_homopolymer_gt10",
                       "lowcmp_SimpleRepeat_imperfecthomopolymer_gt10_slop5",
                       "map_all","map_l250_m0_e0",
                       "notinHG002allgenomespecificandifficult",
                       "notinalldifficultregions",
                       "notinlowcmp_AllRepeats_gt95identity_slop5","segdup")

gg <- ext_trim_df %>% 
    filter(Subtype == "*", Filter == "PASS") %>% 
    filter(Subset %in% strat_of_interest) %>% 
    ggplot() + 
    geom_point(aes(x = METRIC.Recall, 
                   y = METRIC.Precision,
                   alpha = log10(Subset.Size),
                   text = Subset),
               shape = 19, stroke = 0.25) + 
    facet_grid(query_method~Type) + 
    theme_bw() + 
    labs(x = "Recall", y = "Precision")
plotly::ggplotly(gg)
```



__TODO__ Additional figure showing stratifications where the methods perform well and poorly
- stratifications with low recall and precision

Performance metrics are consistent across homopolymer units for homopolymers between 6 and 10 bp but but varies for homopolymers greater than 10 bp. 

```{r fig.cap = "Performance in homopolymers."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"homopolymer")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
        select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90),
          legend.position = "bottom") + 
    labs(x = "Simple Repeat Type", fill = "Callset")
```
DeepVariant worse recall for longer insertions. 
```{r}
ext_trim_df %>% 
    filter(Type == "INDEL") %>% 
    filter(Filter == "PASS") %>% 
    filter(!str_detect(Subtype, "C")) %>% 
    filter(str_detect(Subset,"homopolymer"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    select(Subtype, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Subtype, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA", Metric != "1 - F1_Score") %>% 
    ggplot() + geom_bar(aes(x = Subtype, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Metric~Subset, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90),
          legend.position = "bottom") + 
    labs(x = "Indel Subtype", fill = "Callset")
```




## Manual Curration 

__TODO__ Randomly sampling variants

```{r}
# 
# deep_vcf <- readVcf("data/happy_output/results/result_2.vcf.gz", 
#                     genome = genome_obj)
# 
# get_var_df <- function(vcffile){
#     require(VariantAnnotation)
#     require(BSgenome.Hsapiens.1000genomes.hs37d5)
#     
#     ## Read VCF
#     vcf <- readVcf(vcffile, genome = "BSgenome.Hsapiens.1000genomes.hs37d5")
#     
#     ## Generate tidy data frame
#     
#     
# }
```

```{r}
# ## Truth table classification
# tt_class_df <- geno(gatk_vcf)[['BD']] %>% 
#     as.data.frame() %>% 
#     rownames_to_column(var = "variant") %>% 
#     filter(QUERY == "FP" | TRUTH == "FN") %>% 
#     as_tibble()
# 
# ## Variant type
# var_type_df <- geno(gatk_vcf)[['BVT']] %>% 
#     as.data.frame() %>% 
#     rownames_to_column(var = "variant") %>% 
#     filter(variant %in% tt_class_df$variant)
# 
# ## Filter 
# var_filter_df <- data_frame(
#     variant = rownames(gatk_vcf), 
#     FILTER = rowRanges(gatk_vcf)$FILTER) %>% 
#     filter(variant %in% tt_class_df$variant)
# 
# ## Stratifications
# ## Need to figure out how to handle stratifications
# # region_df <- 
#     # info(head(gatk_vcf))[['Regions']] %>% 
#     # as.data.frame() #%>%
#     # rownames_to_column(var = "variant") # %>% 
#     # filter(variant %in% tt_class_df$variant)
# region_df
# ## Combine data frames
# var_df <- bind_cols(list(tt_class_df , 
#                          var_type_df,
#                          var_filter_df, 
#                          region_df))
```



## Benchmark Region Expansion

# Conclusions


## Exploratory Analysis
### Metric Distributions
```{r message = FALSE, warning = FALSE, fig.cap = "Metric value distributions for all stratifications."}
gatk_metric_df %>% 
    filter(Metric != "METRIC.Frac_NA") %>%
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    ggplot() + 
    geom_histogram(aes(x = Value)) + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw()
```



### Precision-Recall Relationship
Relationships between precision and recall across strata
```{r message = FALSE, fig.cap = "Relationship between precision and recall for all stratifications, excluding different repeat units."}
gg <- ext_trim_df %>% 
    filter(!str_detect(Subset, "unit=")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    filter(Subtype == "*", Filter == "PASS") %>%
    ggplot() + 
    geom_point(aes(x = METRIC.Recall, 
                   y = METRIC.Precision, 
                   alpha = log10(Subset.Size),
                   text = Subset),
               shape = 19, stroke = 0.25) + 
    facet_wrap(query_method~Type, scales = "free") + 
    theme_bw() + 
    labs(x = "Recall", y = "Precision")
plotly::ggplotly(gg)
```


### Difficult Regions

#### GC Impact
GATK4 performance impacted by changes in GC more than DeepVariant. 
Trends in performance relative GC are potentially confounded with low complexity regions.  


```{r fig.cap = "Benchmarking metrics by GC content for 100 bp windows with 50 bp slop." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"gc")) %>% 
    mutate(Subset = str_remove(Subset, "gc_l100_gc")) %>% 
    mutate(Subset = str_remove(Subset, "_slop50")) %>% 
    mutate(Subset = str_replace(Subset, "or", " or ")) %>%
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "lt", "<")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    mutate(Subset = fct_relevel(Subset, 
                                c(">85","<25 or >65", "<30 or >55"), 
                                after = Inf)) %>% 
    select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5,
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90), 
          legend.position = "bottom") +
    labs(x = "% GC for 100 bp windows", fill = "Callset")
```

#### Low Complexity 

##### Simple Repeats
__TODO__ Fix order

```{r fig.cap = "Benchmarking metrics for different types of simple repeats." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"lowcmp_SimpleRepeat"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
        select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90),
          legend.position = "bottom") + 
    labs(x = "Simple Repeat Type", fill = "Callset")
```

##### TRDB
__Q__ Difference between all_gt95identity_merged and all_merged  
__TODO__ Fix order  

```{r fig.cap = "Benchmarking metrics for TRDB." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"Human_Full_Genome"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>%
    mutate(Subset = str_remove(Subset, "lowcmp_Human_Full_Genome_TRDB_hg19_150331_")) %>%
    mutate(Subset = str_remove(Subset, "_gt95identity_merged")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>%
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "lt", " <")) %>%
    mutate(Subset = str_replace(Subset, "gt", " >")) %>%
        select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Tandem Repeat Type", fill = "Callset")
```

##### All Repeats
__TODO__ Fix order, add figure caption describing stratifications

```{r fig.cap = "Benchmarking metrics for all repeats."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"AllRepeats"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>%
    mutate(Subset = str_remove(Subset, "lowcmp_AllRepeats_")) %>%
    mutate(Subset = str_remove(Subset, "_gt95identity_merged")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>%
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "lt", "<")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Repeat Type", fill = "Callset")
```


##### Segmental Duplications
```{r fig.cap = "Segmental duplications"}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"segdup")) %>% 
    select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Segmental Duplication Category", fill = "Callset")
```

##### Mappping
__TODO__ Fix order and clean-up names  
__Q__ What is Siren  
```{r fig.cap = "Mapping." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"map"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>%
    mutate(Subset = str_remove(Subset, "map_l")) %>%
    mutate(Subset = str_replace_all(Subset, "_", " ")) %>%
    select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) +
    labs(x = "Mapping Category", fill = "Callset")
```

### In/ Not-In Comparisons
```{r fig.cap = "Benchmarking metrics for in/not-in comparisons"}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(Subset %in% c("notinfunc_cds","func_cds",
                         "notinsegdupall", "segdupall",
                         "notinalldifficultregions", "alldifficultregions")) %>% 
    mutate(Subset = fct_relevel(Subset, "func_cds", after = 2)) %>%
    mutate(Subset = fct_relevel(Subset, "notinsegdupall", after = Inf)) %>%
    select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Stratification", fill = "Callset")
```

### Performs well
```{r}
ext_trim_df %>% 
    # filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>%
    filter(!str_detect(Subset, "unit")) %>% 
    filter(METRIC.Recall >= 0.999, 
           METRIC.Precision > 0.98)
```



# Session Information
## System Information
```{r}
s_info <- devtools::session_info(include_base = FALSE)
pander::pander(s_info$platform)
```

## Package Versions
```{r}
s_info$packages %>% filter(`*` == "*") %>% select(-`*`) %>%
    knitr::kable(booktabs = TRUE)
```