---
title: "HG002 PacBio CCS Small Variants"
author: "Nate Olson"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r load_packages, message = FALSE, echo = FALSE}
library(VariantAnnotation)
library(BSgenome.Hsapiens.1000genomes.hs37d5)
library(happyR)
library(tidyverse)
library(googlesheets)
```

# Overview
Analysis of small variant callsets generated from PacBio CCS data. 
The long read - low error rate data can potentially correct mistakes in and expand the HG002 benchmark set.  

* datasets - PacBio CCS 15kb libraries for HG002  
* variant callsets - Three SNV+indel callsets: GATK4, DeepVariant, and DeepVariant with haplotype information. Variant calls were made against the GRCh37 reference.  
* benchmarking - Callsets benchmarked against the HG002 V3.3.2 benchmark callset using the precisionFDA app, vcfeval + hap.py with GA4GH custom stratifications  

# Approach/ Methods
## Benchmarking 
- Stratified comparison to our current small variant benchmark (v3.3.2).   
- Identify where callsets perform well and poorly  
- Identify where callsets perform well with a high fraction of calls in a stratification outside the benchmark regions.   

## Manual Curation
A number of false positive and negative sites were randomly sampled to further evaluate the potential for using CCS callsets to correct errors in benchmark sets.
A total of 60 variants were randomly sampled for manual curation.  

- 5 from each set  
    - FP and FN  
    - SNP and Indels  
    - Target categories and other categories  

Along with 5 SNP and 5 Indel FP allele match errors.  

Target categories - lowcmp_AllRepeats_gt95identity_slop5 and lowcmp_SimpleRepeat_imperfecthomopolymer_gt10_slop5

<!-- 
Especially for indels, most of the errors are in homopolymers, so it’s helpful to segregate these. 
For SNPs, it might also be useful to select 10 random from the most difficult to map regions (map_l250_m0_e0), unless most of the non-homopolymer SNPs are in these regions. 
Definitely let me know if your look at the data suggests a different strategy though! I can help with the manual curation when I return from AMP. 
-->

## Benchmark Region Expansion Estimate
Estimating of how many extra variants and regions CCS might help add to the GIAB benchmark. Perhaps, regions with good MQ and coverage excluding homopolymers, and the CCS variants in these regions.  

<!-- - Need to work out approach -->
<!--     - check out gvcf for good MQ and coverage   -->

# Results 

## Benchmarking  

<!-- - Subsets to look at  -->
<!--     - `*`   -->
<!-- 	- HG002allgenomespecific   -->
<!-- 	- HG002allgenomespecificandifficult   -->
<!-- 	- alldifficultregions   -->
<!-- 	- decoy   -->
<!-- 	- lowcmp_AllRepeats_51to200bp_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_AllRepeats_gt200bp_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_AllRepeats_gt95identity_slop5   -->
<!-- 	- lowcmp_AllRepeats_lt51bpTRs_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_AllRepeats_lt51bp_gt95identity_merged_slop5   -->
<!-- 	- lowcmp_SimpleRepeat_homopolymer_6to10   -->
<!-- 	- lowcmp_SimpleRepeat_homopolymer_gt10   -->
<!-- 	- lowcmp_SimpleRepeat_imperfecthomopolymer_gt10_slop5   -->
<!-- 	- map_all   -->
<!-- 	- map_l250_m0_e0   -->
<!-- 	- notinHG002allgenomespecificandifficult   -->
<!-- 	- notinalldifficultregions   -->
<!-- 	- notinlowcmp_AllRepeats_gt95identity_slop5   -->
<!-- 	- segdup   -->

```{r message = FALSE}
## Loading data
hap_list <- c(paste0("data/happy_output/results/result_", c(1,2)),
              "data/happy_output_deepVarHap/results/result_1") %>%
    map(read_happy) %>% 
    set_names(c("GATK4","DeepVariant", "DeepVarHap"))

## Creating a tidy data frame
extend_df <- hap_list %>% 
    map("extended") %>% 
    bind_rows(.id = "query_method")

ext_trim_df <- extend_df %>% 
    ## Excluding non-HG002 genome specific stratifications
    filter(!str_detect(Subset, "HG00[1,3,4,5]")) %>% 
    ## excluding strat with < 1000 obs - lack of statistical power
    filter(Subset.Size > 1000) %>% 
    ## Subset must have at least one variant in region
    filter(Subset.IS_CONF.Size > 0)

## For initial set of plots
gatk_ext_df <-  filter(ext_trim_df, 
                       query_method == "GATK4", 
                       Subtype == "*")

gatk_metric_df <- gatk_ext_df %>% 
    group_by(Type, Subset) %>% 
    dplyr::select(contains("METRIC")) %>% 
    gather("Metric", "Value", -Type, -Subset) 

```


Overall high precision and recall for SNPs, and indels not in difficult regions. 
DeepVariant performs consistently better than GATK4. 
The high frac_NA for SNPs in all difficult regions along with relatively high precision and recall indicates that using this in the GIAB integration pipeline may result in expanded benchmark regions.  
- Difficult regions with low indel recall and precision, specifically homopolymers and imperfect homopolymers 

Other observations  
- Lower than expected precision for mappable regions - potentially errors in benchmark set. Also interesting that precision is lower for m and e 0 than 1 or 2. Could this be related to Alu elements?


```{r}
## Do we have all difficult regions not in HG002allgenomespecific
high_level_strat <- c("*",
                       "alldifficultregions",
                       "notinHG002allgenomespecificandifficult",
                       "notinalldifficultregions")

ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(Subset %in% high_level_strat) %>% 
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -45, hjust = 0))
```

```{r}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Subset %in% high_level_strat) %>% 
    filter(Filter == "PASS") %>% 
    rename(Callset = query_method,
            Recall = METRIC.Recall, 
           Precision = METRIC.Precision, 
           F1 = METRIC.F1_Score, 
           Frac_NA = METRIC.Frac_NA) %>%
    dplyr::select(Callset, Type, Subset, Recall, Precision, F1, Frac_NA, Subset.Size) %>% 
    arrange(Type, Subset) %>% 
    knitr::kable(digits = 3, 
                 caption = "Metrics for high level stratifications.")
```

high precision and recall
```{r, fig.cap = "GATK4 and DeepVariant precision v. recall for a subset of the stratifications."}
strat_of_interest <- c("*","HG002allgenomespecific",
                       "HG002allgenomespecificandifficult",
                       "alldifficultregions","decoy", 
                     "lowcmp_AllRepeats_51to200bp_gt95identity_merged_slop5",
                       "lowcmp_AllRepeats_gt200bp_gt95identity_merged_slop5",
                       "lowcmp_AllRepeats_gt95identity_slop5",
                    "lowcmp_AllRepeats_lt51bpTRs_gt95identity_merged_slop5",
                       "lowcmp_AllRepeats_lt51bp_gt95identity_merged_slop5",
                       "lowcmp_SimpleRepeat_homopolymer_6to10",
                       "lowcmp_SimpleRepeat_homopolymer_gt10",
                       "lowcmp_SimpleRepeat_imperfecthomopolymer_gt10_slop5",
                       "map_all","map_l250_m0_e0",
                       "notinHG002allgenomespecificandifficult",
                       "notinalldifficultregions",
                       "notinlowcmp_AllRepeats_gt95identity_slop5","segdup")

gg <- ext_trim_df %>% 
    filter(Subtype == "*", Filter == "PASS") %>% 
    filter(Subset %in% strat_of_interest) %>% 
    ggplot() + 
    geom_point(aes(x = METRIC.Recall, 
                   y = METRIC.Precision,
                   alpha = log10(Subset.Size),
                   text = Subset),
               shape = 19, stroke = 0.25) + 
    facet_grid(query_method~Type) + 
    theme_bw() + 
    labs(x = "Recall", y = "Precision")
plotly::ggplotly(gg)
```

```{r}
ext_trim_df %>% 
    filter(Subtype == "*", Filter == "PASS") %>% 
    filter(Subset %in% strat_of_interest) %>% 
    dplyr::select(query_method, Type, Subset, contains("METRIC")) %>% 
    DT::datatable(caption = "The three query methods (variant callsets) performance metrics for high level stratifications and stratifications with know poor performance, e.g. homopolymers.")
```


```{r}
ext_trim_df %>% 
    filter(Type == "INDEL", Filter == "PASS") %>% 
    filter(Subset %in% strat_of_interest) %>% 
    filter(TRUTH.TOTAL > 0)%>% 
    dplyr::select(query_method, Type, Subtype,Subset, contains("METRIC"), TRUTH.TOTAL) %>%
     DT::datatable()
```

__TODO__ Additional figure showing stratifications where the methods perform well and poorly
- stratifications with low recall and precision

Performance metrics are consistent across homopolymer units for homopolymers between 6 and 10 bp but but varies for homopolymers greater than 10 bp. 

```{r fig.cap = "Performance in homopolymers."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"homopolymer")) %>% 
    filter(!str_detect(Subset,"_unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
        dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = query_method, y = Value, fill = Subset), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(legend.position = "bottom") + 
    labs(x = "Simple Repeat Type", fill = "Homopolymer Type")
```


```{r fig.cap = "DeepVariant performance metrics for homopolymers by unit."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"homopolymer")) %>% 
    filter(str_detect(Subset,"_unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    separate(Subset, c("Size", "Unit"), sep = "_") %>% 
    mutate(Unit = str_remove(Unit, "unit=")) %>% 
    mutate(Size = str_remove(Size, "homopolymer ")) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    filter(query_method == "DeepVariant") %>% 
    ggplot() + geom_bar(aes(x = Size, y = Value, fill = Unit), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(legend.position = "bottom") + 
    labs(x = "Simple Repeat Type", fill = "Callset")
```

```{r fig.cap = "DeepVariant with haplotype informaton performance metrics for homopolymers by unit."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"homopolymer")) %>% 
    filter(str_detect(Subset,"_unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    separate(Subset, c("Size", "Unit"), sep = "_") %>% 
    mutate(Unit = str_remove(Unit, "unit=")) %>% 
    mutate(Size = str_remove(Size, "homopolymer ")) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    filter(query_method == "DeepVarHap") %>% 
    ggplot() + geom_bar(aes(x = Size, y = Value, fill = Unit), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(legend.position = "bottom") + 
    labs(x = "Simple Repeat Type", fill = "Callset")
```

```{r fig.cap = "GATK4 performance metrics for homopolymers by unit."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"homopolymer")) %>% 
    filter(str_detect(Subset,"_unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    separate(Subset, c("Size", "Unit"), sep = "_") %>% 
    mutate(Unit = str_remove(Unit, "unit=")) %>% 
    mutate(Size = str_remove(Size, "homopolymer ")) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    filter(query_method == "GATK4") %>% 
    mutate(Unit = fct_relevel(Unit, c("A","T","C","G"))) %>% 
    ggplot() + geom_bar(aes(x = Size, y = Value, fill = Unit), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(legend.position = "bottom") + 
    labs(x = "Simple Repeat Type", fill = "Unit")
```


DeepVariant worse recall for longer insertions. 
```{r}
ext_trim_df %>% 
    filter(Type == "INDEL") %>% 
    filter(Filter == "PASS") %>% 
    filter(!str_detect(Subtype, "C")) %>% 
    filter(str_detect(Subset,"homopolymer"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    dplyr::select(Subtype, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Subtype, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA", Metric != "1 - F1_Score") %>% 
    ggplot() + geom_bar(aes(x = Subtype, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Metric~Subset, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90),
          legend.position = "bottom") + 
    labs(x = "Indel Subtype", fill = "Callset")
```




## Manual Curration 

```{r message = FALSE}
get_var_df <- function(vcffile){
    ## Read VCF
    vcf <- readVcf(vcffile, genome = "BSgenome.Hsapiens.1000genomes.hs37d5")

    ## Generate tidy data frame
    ## Truth table classification
    tt_class_df <- geno(vcf)[['BD']] %>%
        as.data.frame() %>%
        rownames_to_column(var = "variant") %>%
        filter(QUERY == "FP" | TRUTH == "FN") %>%
        as_tibble()

    ## Subsetting vcf for FP and FN
    fpfn_positions <- str_remove(tt_class_df$variant, "_.*")
    vcf_positions <- rownames(vcf) %>% str_remove("_.*")
    fpfn_vcf <- vcf[vcf_positions %in% fpfn_positions ]
    
    ## Truth table classifications for Fp and Fn positions
     tt_fpfn_df <- geno(fpfn_vcf )[['BD']] %>%
        as.data.frame() %>%
        rownames_to_column(var = "variant") %>%
        as_tibble()
     
    ## Benchmarking Type 
    bk_type_df <- geno(fpfn_vcf)[['BK']] %>% 
        as.data.frame() %>%
        rownames_to_column(var = "variant") %>% 
        dplyr::rename(Q.bk = QUERY, T.bk = TRUTH) %>%
        as_tibble() %>% 
        left_join(tt_class_df, by = c("variant" = "variant"))
    
    ## Genotype 
    gt_type_df <- geno(fpfn_vcf)[['GT']] %>% 
        as.data.frame() %>%
        rownames_to_column(var = "variant") %>%
        dplyr::rename(Q.gt = QUERY, T.gt = TRUTH) %>% 
        as_tibble() %>% 
        left_join(bk_type_df)
    
    ## Variant type
    fpfn_df <- geno(fpfn_vcf)[['BVT']] %>%
        as.data.frame() %>%
        rownames_to_column(var = "variant") %>%
        dplyr::rename(Q.Var = QUERY, T.Var = TRUTH) %>%
        left_join(gt_type_df)

    ## Combining into single data frame
    fpfn_df %>% add_column(FILTER = rowRanges(fpfn_vcf)$FILTER,
                           Regions = as.list(info(fpfn_vcf)[['Regions']]))
}

annotate_var_df <-  function(var_df){
    var_anno_df <- var_df %>% 
        mutate(allrepeats = map_lgl(Regions, ~("lowcmp_AllRepeats_gt95identity_slop5" %in% .)),
               imphomo = map_lgl(Regions, ~("lowcmp_SimpleRepeat_imperfecthomopolymer_gt10_slop5" %in% .)),
               target_cat = if_else(allrepeats + imphomo == 0, "non-target","target")) %>%
        dplyr::select(-allrepeats, -imphomo)

    mutate(var_anno_df, 
           var_cat = case_when( 
               (Q.bk == "am" | T.bk == "am") & (Q.Var == "INDEL" | T.Var == "INDEL") ~ "AM.INDEL",
               (Q.bk == "am" | T.bk == "am") & (Q.Var == "SNP" | T.Var == "SNP") ~ "AM.SNP",
               # Q.bk == "am" | T.bk == "am" ~ "AM",
               QUERY == "FP" & Q.Var == "INDEL" ~ "FP.INDEL",
               QUERY == "FP" & Q.Var == "SNP" ~ "FP.SNP",
               QUERY == "." & T.Var == "INDEL" ~ "FN.INDEL",
               QUERY == "." & T.Var == "SNP" ~ "FN.SNP",
               TRUE ~ "other")
           )
}
```


```{r}
## Getting random subset
anno_df <- list(gatk = "data/happy_output/results/result_1.vcf.gz",
                deep = "data/happy_output/results/result_2.vcf.gz",
                deepHap = "data/happy_output_deepVarHap/results/result_1.vcf.gz") %>% 
    map(get_var_df) %>% 
    map_dfr(annotate_var_df, .id = "callset") %>% 
    # Cleaner variant ids
    mutate(CHROM = str_extract(variant, ".*(?=:)"), 
           POS = str_extract(variant, "(?<=:).*(?=_)"), 
           ALT = str_extract(variant, "(?<=_).*"))
    
set.seed(531)
var_random_df <- anno_df %>% 
    dplyr::filter(var_cat != "other") %>% 
    mutate(CHROM = str_extract(variant, ".*(?=:)")) %>% 
    mutate(POS = str_extract(variant, "(?<=:).*(?=_)")) %>% 
    ## Random subsetting to include multiple variant groups
    group_by(callset, var_cat, target_cat) %>%
    sample_n(size = 5) %>%
    ungroup() %>% 
    ## Includes ALTs at the same position
    dplyr::select(callset, CHROM, POS) %>%
    left_join(anno_df)

deepvar_random_df <- var_random_df %>% 
    filter(callset == "deepHap") %>%
    mutate(variant = str_remove(variant, ".*_")) %>% 
    dplyr::rename(VAR = variant) %>% 
    dplyr::select(-Regions, -ALT) %>% 
    select(callset, target_cat, var_cat, CHROM, POS, everything())
```


Total number of variants in each variant category for target and non-target regions.
```{r}
anno_df %>%
    dplyr::select(callset, CHROM, POS, target_cat, var_cat) %>% 
    distinct() %>% 
    group_by(callset, target_cat, var_cat) %>% 
    summarise(count = n()) %>% 
    spread(target_cat, count)
```

```{r fig.height = 12, fig.width = 12, fig.cap = "Heatmap with stratifications for the random variant subset selected for manual curation."}
var_region_heatmap_df <- var_random_df %>% 
    filter(callset == "deepHap") %>%
    dplyr::select(variant, var_cat, target_cat, Regions) %>% 
    mutate(region_df = map(Regions, as_data_frame)) %>% 
    dplyr::select(-Regions) %>% 
    unnest()
var_region_heatmap_df %>% 
    filter(!str_detect(value, "HG00[1,3,4,5]"), value !="CONF") %>%
    mutate(CHROM_POS = str_extract(variant, ".*(?=_)")) %>%
    mutate(ALT = str_extract(variant, "(?<=_).*")) %>%
    ggplot() + 
    geom_raster(aes(y = value, x = CHROM_POS, fill = target_cat)) + 
    theme(axis.text.x = element_text(angle = -90)) + 
    facet_wrap(~var_cat, nrow = 1, scales = "free_x")
```


### Manual curation spreadsheets
```{r}
tmp_tsv <- tempfile(fileext = ".tsv")
deepvar_random_df %>%
    add_column(PacBio = "", GIAB = "", Notes = "") %>%
    write_tsv(path = tmp_tsv)

gs_upload(file = tmp_tsv, sheet_title = "hg002-ccs-deepvar-curate-JZ")
gs_upload(file = tmp_tsv, sheet_title = "hg002-ccs-deepvar-curate-JM")
gs_upload(file = tmp_tsv, sheet_title = "hg002-ccs-deepvar-curate-NDO")
```


## Benchmark Region Expansion
__TODO__

# Conclusions
__TODO__

# Exploratory Analysis
## Metric Distributions
```{r message = FALSE, warning = FALSE, fig.cap = "Metric value distributions for all stratifications."}
gatk_metric_df %>% 
    filter(Metric != "METRIC.Frac_NA") %>%
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    ggplot() + 
    geom_histogram(aes(x = Value)) + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw()
```



## Precision-Recall Relationship
```{r message = FALSE, fig.cap = "Relationship between precision and recall for all stratifications, excluding different repeat units."}
gg <- ext_trim_df %>% 
    filter(!str_detect(Subset, "unit=")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    filter(Subtype == "*", Filter == "PASS") %>%
    ggplot() + 
    geom_point(aes(x = METRIC.Recall, 
                   y = METRIC.Precision, 
                   alpha = log10(Subset.Size),
                   text = Subset),
               shape = 19, stroke = 0.25) + 
    facet_wrap(query_method~Type, scales = "free") + 
    theme_bw() + 
    labs(x = "Recall", y = "Precision")
plotly::ggplotly(gg)
```


## GC Impact
GATK4 performance impacted by changes in GC more than DeepVariant and DeepVariant with haplotype information. 
Trends in performance relative GC are potentially confounded with low complexity regions.  


```{r fig.cap = "Benchmarking metrics by GC content for 100 bp windows with 50 bp slop." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"gc")) %>% 
    mutate(Subset = str_remove(Subset, "gc_l100_gc")) %>% 
    mutate(Subset = str_remove(Subset, "_slop50")) %>% 
    mutate(Subset = str_replace(Subset, "or", " or ")) %>%
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "lt", "<")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    mutate(Subset = fct_relevel(Subset, 
                                c(">85","<25 or >65", "<30 or >55"), 
                                after = Inf)) %>% 
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5,
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90), 
          legend.position = "bottom") +
    labs(x = "% GC for 100 bp windows", fill = "Callset")
```

## Difficult Regions

### Low Complexity 

#### Simple Repeats
__TODO__ Fix order

```{r fig.cap = "Benchmarking metrics for different simple repeat stratifications." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"lowcmp_SimpleRepeat"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>% 
    mutate(Subset = str_remove(Subset, "lowcmp_SimpleRepeat_")) %>%
    mutate(Subset = str_remove(Subset, "_slop5")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>% 
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
        dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90),
          legend.position = "bottom") + 
    labs(x = "Simple Repeat Type", fill = "Callset")
```

#### TRDB
__Q__ Difference between all_gt95identity_merged and all_merged  
__TODO__ Fix order  

```{r fig.cap = "Benchmarking metrics for TRDB." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"Human_Full_Genome"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>%
    mutate(Subset = str_remove(Subset, "lowcmp_Human_Full_Genome_TRDB_hg19_150331_")) %>%
    mutate(Subset = str_remove(Subset, "_gt95identity_merged")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>%
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "lt", " <")) %>%
    mutate(Subset = str_replace(Subset, "gt", " >")) %>%
        dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Tandem Repeat Type", fill = "Callset")
```

#### All Repeats
__TODO__ Fix order

```{r fig.cap = "Benchmarking metrics for all repeat stratifications."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"AllRepeats"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>%
    mutate(Subset = str_remove(Subset, "lowcmp_AllRepeats_")) %>%
    mutate(Subset = str_remove(Subset, "_gt95identity_merged")) %>%
    mutate(Subset = str_replace(Subset, "_", " ")) %>%
    mutate(Subset = str_replace(Subset, "to", " to ")) %>%
    mutate(Subset = str_replace(Subset, "lt", "<")) %>%
    mutate(Subset = str_replace(Subset, "gt", ">")) %>%
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Repeat Type", fill = "Callset")
```


### Segmental Duplications
```{r fig.cap = "Performance metrics for segmental duplication stratifications."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"segdup")) %>% 
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Segmental Duplication Category", fill = "Callset")
```

### Mappping
__TODO__ Fix order and clean-up names  
__Q__ What is Siren  
```{r}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(Subset %in% c("map_all", "map_siren", "notinmap_all")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>%
    mutate(Subset = str_remove(Subset, "map_l")) %>%
    mutate(Subset = str_replace_all(Subset, "_", " ")) %>%
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) +
    labs(x = "Mapping Category", fill = "Callset")
```


```{r fig.cap = "Performance metrics for mappability stratifications." }
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(str_detect(Subset,"map"),
           !str_detect(Subset,"unit")) %>% 
    filter(Subset.IS_CONF.Size > 0) %>%
    mutate(Subset = str_remove(Subset, "map_l")) %>%
    mutate(Subset = str_replace_all(Subset, "_", " ")) %>%
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) +
    labs(x = "Mapping Category", fill = "Callset")
```

## In/ Not-In Comparisons
```{r fig.cap = "Benchmarking metrics for in/not-in stratification comparisons."}
ext_trim_df %>% 
    filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>% 
    filter(Subset %in% c("notinfunc_cds","func_cds",
                         "notinsegdupall", "segdupall",
                         "notinalldifficultregions", "alldifficultregions")) %>% 
    mutate(Subset = fct_relevel(Subset, "func_cds", after = 2)) %>%
    mutate(Subset = fct_relevel(Subset, "notinsegdupall", after = Inf)) %>%
    dplyr::select(Type, Subset, query_method, contains("METRIC"), Subset.Size)  %>% 
    gather("Metric","Value", -Type, -Subset, 
           -query_method, -Subset.Size) %>% 
    mutate(Metric = str_remove(Metric, "METRIC.")) %>% 
    mutate(Metric = if_else(Metric != "Frac_NA", paste("1 -", Metric), Metric),
           Value = if_else(Metric != "Frac_NA", 1 - Value, Value)) %>%
    filter(Metric != "Frac_NA") %>% 
    ggplot() + geom_bar(aes(x = Subset, y = Value, fill = query_method), 
                        color = "grey40", width = 0.5, 
                        position = "dodge", stat = "identity") + 
    facet_grid(Type~Metric, scales = "free") + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = -90)) + 
    labs(x = "Stratification", fill = "Callset")
```

## Performs well
Statifications with high precision and recall  
```{r}
ext_trim_df %>% 
    # filter(Subtype == "*") %>% 
    filter(Filter == "PASS") %>%
    filter(!str_detect(Subset, "unit")) %>% 
    filter(METRIC.Recall >= 0.999, 
           METRIC.Precision > 0.98) %>% 
    dplyr::select(query_method, Type, Subtype, Subset, 
           contains("METRIC"), 
           Subset.Size, Subset.IS_CONF.Size)
```



# Session Information
## System Information
```{r}
s_info <- devtools::session_info(include_base = FALSE)
pander::pander(s_info$platform)
```

## Package Versions
```{r}
s_info$packages %>% filter(`*` == "*") %>% dplyr::select(-`*`) %>%
    knitr::kable(booktabs = TRUE)
```